Forecasting the 2024 U.S. Presidential Election: A Poll-of-Polls Approach for Predicting the Outcome

### 2024-11-04 paper/paper.qmd
paper with problem

# Model

@@

The model captures both the fixed effect of poll percentage () and the random effects attributable to differences between pollsters and states.

@@

## Visualization of Results

@@

To visualize the variability in predicted probabilities by state, the following plot shows the average predicted probability of Kamala Harris winning in each state:

@@

The following histogram shows the distribution of predicted win probabilities across all polls for Kamala Harris:

>>>>

# Model
```{r}
# Load libraries and data
library(tidyverse)
library(ggplot2)
library(rstanarm)
library(arrow)

clean_president_polls <- read_parquet("../data/02-analysis_data/cleaned_president_polls.parquet")

clean_president_polls <- clean_president_polls %>%
  mutate(
    weight = numeric_grade * (sample_size / mean(sample_size)),
    pollster = as.factor(pollster),
    state = as.factor(state),
    candidate_label = ifelse(is_harris == 1, "Harris", "Trump")
  )

@@

The model captures both the fixed effect of poll percentage () and the random effects attributable to differences between pollsters and states.
```{r}
# Fit Bayesian model
bayesian_model <- stan_glmer(
  is_harris ~ pct + (1 | pollster) + (1 | state),
  data = clean_president_polls,
  family = binomial(link = "logit"),
  prior = normal(0.5, 0.1, autoscale = TRUE),
  prior_intercept = normal(0.5, 0.1, autoscale = TRUE),
  weights = weight,
  cores = 4,
  adapt_delta = 0.95,
  refresh = 0
)

# Predict probabilities
clean_president_polls <- clean_president_polls %>%
  mutate(
    predicted_prob_harris = posterior_predict(bayesian_model, newdata = clean_president_polls, type = "response") %>% colMeans(),
    winner_harris = ifelse(predicted_prob_harris > 0.5, 1, 0)
  )

# Plot distribution of predicted win probabilities for Harris
clean_president_polls %>%
  ggplot(aes(x = predicted_prob_harris)) +
  geom_histogram(binwidth = 0.05, fill = "blue", color = "black", alpha = 0.7) +
  labs(title = "Distribution of Predicted Win Probabilities for Kamala Harris",
       x = "Predicted Win Probability",
       y = "Count") +
  theme_minimal()

@@

## Visualization of Results
```{r}
# Predict probabilities
clean_president_polls <- clean_president_polls %>%
  mutate(
    predicted_prob_harris = posterior_predict(bayesian_model, newdata = clean_president_polls, type = "response") %>% colMeans(),
    winner_harris = ifelse(predicted_prob_harris > 0.5, 1, 0)
  )

# Calculate average predicted probability for Harris by state
state_predictions <- clean_president_polls %>%
  group_by(state) %>%
  summarize(
    avg_predicted_prob_harris = mean(predicted_prob_harris),
    state_winner = ifelse(avg_predicted_prob_harris > 0.5, "Harris", "Trump")
  )

@@

To visualize the variability in predicted probabilities by state, the following plot shows the average predicted probability of Kamala Harris winning in each state:
```{r}
#| echo: false
# Plot predicted probability of Harris winning by state
state_predictions %>%
  ggplot(aes(x = reorder(state, avg_predicted_prob_harris), y = avg_predicted_prob_harris, fill = state_winner)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Predicted Probability of Harris Winning by State",
       x = "State",
       y = "Average Predicted Probability",
       fill = "Predicted Winner") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = rel(0.6)))

@@

The following histogram shows the distribution of predicted win probabilities across all polls for Kamala Harris:
```{r}
# Predict probabilities and determine winners
clean_president_polls <- clean_president_polls %>%
  mutate(
    predicted_prob_harris = posterior_predict(bayesian_model, newdata = clean_president_polls, type = "response") %>% colMeans(),
    winner_harris = ifelse(predicted_prob_harris > 0.5, 1, 0)
  )
